### 2016-06-25 project/paper/sections/master_project.tex
* refactored server / client * refined data stream * added data stream mapping

\label{chap:master_project}
\begin{itemize}
\item InfoInside, WVK, Tag der offenen Tür
\item Latex Vorlage
\item Doku / Einführung Fahrsimulator, CANoe / OpenDS support
\item EEG einschicken
\end{itemize}